---
title: "Lab 01 - Hello R"
author: Eric!
date: 12.30.23
output: 
  html_document: 
    fig_height: 5.33
    fig_width: 12
    theme: sandstone
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

Based on the help file, how many rows and how many columns does the datasaurus_dozen file have? What are the variables included in the data frame? Add your responses to your lab report.

According to the help, there are 1846 rows and 3 variables. I was confused by this at first, because if there are lots of data sets, how can there just be 3 variables? But I'm now thinking of it that there are 3 *columns* where 2 of the columns are essentially *x* and *y*, and x and y take on different meanings in the different data sets (e.g., for dino, dots, etc). Is that accurate?

The three variables / columns are:  
(1) which data set the data are from
(2) x-values
(3) y-values

The different data sets are:

 1 away      
 2 bullseye  
 3 circle    
 4 dino      
 5 dots      
 6 h_lines   
 7 high_lines
 8 slant_down
 9 slant_up  
10 star      
11 v_lines   
12 wide_lines
13 x_shape  


### Exercise 2

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

The graph looks like a dinosaur. Note if you change the commands to:

ggplot(data = dino_data, mapping = aes(x = y, y = x)) +
  geom_point()

the dinosaur is lying down!

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

The correlation is -0.0645

### Exercise 3

Plot y vs. x for the star dataset. You can (and should) reuse code we introduced above, just replace the dataset name with the desired dataset. Then, calculate the correlation coefficient between x and y for this dataset. How does this value compare to the r of dino?

Add code and narrative as needed.
Note that the R chunks are labeled with `plot-star` and `cor-star` to provide spaces to place the code for plotting and calculating the correlation coefficient.
To finish, clean up the narrative by removing these instructions.



```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

The graph looks like a star.

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

r = -.0630

Basically the same correlation from before. So just based on the linear correlations, the associations between x and y are identical. Hence a good example of why linear relationships don't tell the whole story! :)

### Exercise 4

Add code and narrative as needed.
Note that two R chunks are given but they are not labeled.
Use the convention from above to name them appropriately.


```{r}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r}
circle_data %>%
  summarize(r = cor(x, y))
```

The correlation is -.0683 ... again, the same as in dino.

### Exercise 5

Add code and narrative as needed.

To add R chunks either type out the backticks, curly braces, and the letter `r` or use the Insert chunk button above, green C+.

```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point()+
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")

datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```

Note when I run the above command it shows the correlations, but when I knitted it the correlations did not appear.
Which makes me think the second variable (i.e., the correlation) isn't showing up. Shouldn't run and knit produce the same results?

